Individual and Differential Harm in Redistricting

### 2022-04-08 paper/paper.Rmd
cleanup boston process; rasterize pairs points

The expected number of Democratic seats and the differential partisan harm measures identify each gerrymanders as an outlier in the correct direction (two-sided $p=0.0002$ for expected seats), and as expected from Proposition \ref{prop:fairness}, these two measures are highly correlated.
The efficiency gap and declination measures also correctly identify the gerrymanders ($p=0.0002$), though they are much more discrete, and some nonlinearity is evident for declination.
In contrast, the partisan bias and mean-median measures incorrectly classify the Republican gerrymander as a fair plan and a Democratic-leaning outlier, respectively ($p=0.099$ and $p=0.0022$).
These measures are weakly correlated with the expected number of Democratic seats, and show substantial variation even for fixed values of total seats won at baseline.
There is also disagreement between the measures about the natural bias of districting plans in New Jersey: the median sampled plan has a partisan bias of $0.047$, indicating a natural Republican bias, but the median mean-median score is just $0.0069$, and the median differential partisan harm is $-0.13$, indicating a natural Democratic bias.
Which is correct?
Democrats win 8.47 seats on average across the simulations, which is more than the 6.79 seats that would be expected under a completely proportional outcome.

>>>>

Differential partisan harm identifies each gerrymander as such, with a large negative value for the Democratic gerrymander and a large positive value for the Republican gerrymander.
These values are relatively directly interpretable; for example, the differential harm score of $-0.22$ means that an average Democratic voter is 22% less likely to be misrepresented than their average Republican counterpartisan.
The efficiency gap and declination measures also correctly identify the gerrymanders, though they are much more discrete, and declination gives a much larger score in absolute value to the Democratic plan than to the Republican.
In contrast, the partisan bias and mean-median measures incorrectly classify the Republican gerrymander as a fair plan and a Democratic-leaning outlier.
These measures are weakly correlated with the expected number of Democratic seats, and show substantial variation even for fixed values of total seats won at baseline.

Considering the distribution of these statistics across the distribution of sampled plans doesn't help much, either.
The expected number of Democratic seats naturally identifies the gerrymanders (two-sided $p=0.0002$) and the direction of bias, and as expected from Proposition \ref{prop:fairness} this measure is highly correlated with differential partisan harm.
The gerrymanders are also an outlier under the efficiency gap and declination measures also (both $p=0.0002$).
While the gerrymanders are both an outlier for the mean-median score, only the Democratic plan is an outlier according to the partisan bias measure ($p=0.0022$, versus $p=0.099$ for the Republican plan).

There is also disagreement between the measures about the natural bias of districting plans in New Jersey: the median sampled plan has a partisan bias of $0.047$, indicating a natural Republican bias, but the median mean-median score is just $0.0069$, and the median differential partisan harm is $-0.13$, indicating a natural Democratic bias.
Which is correct?
Democrats win 8.47 seats on average across the simulations, which is more than the 6.79 seats that would be expected under a completely proportional outcome.
